---
title: "Numeric Exploratory Data Analysis"
author: "Siyue Gao"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data

```{r}
library(tidyverse)
```

This code chunk from the course website gets the data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>% # rounds down the date to the bottom month
  select(name, id, everything())
```


## `group_by()`

```{r}
weather_df %>% 
  group_by(name)

weather_df %>% 
  group_by(month)

weather_df %>% 
  group_by(name, month)
```


## `summarise()`

let's group and then summarize!

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarise(
    n_obs = n()
  )

weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    n_obs = n()
  )
```

can we count in other ways?

```{r}
weather_df %>% 
  count(name, month)
```

don't use `table()` (baseR's built-in function)

the output is not a dataframe!

```{r, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

you can have multiple summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarise(
    n_obs = n(),
    n_dist = n_distinct()
  )

weather_df %>% 
  group_by(month) %>% 
  summarise(
    n_obs = n(),
    n_dist = n_distinct(month)
  )

weather_df %>% 
  group_by(month) %>% 
  summarise(
    n_obs = n(),
    n_dist = n_distinct(date)
  )
```


## a brief aside on 2x2 tables

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "" # what will happen by default
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  pivot_wider( # turn into a 2x2 table by `pivot_wider()`
    names_from = "cold",
    values_from = "n_obs"
  )
```

other ways to make a 2x2 table by a `janitor()` package, quicker than `pivot_wider()`

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```


## General summaries

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    n_obs = n(),
    mean_tmax = mean(tmax)
  )

weather_df %>% 
  filter(is.na(tmax))

weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    sd_prcp = sd(prcp),
    median_tmax = median(tmax, na.rm = TRUE)
  )
```


```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(across(prcp:tmin, mean))

weather_df %>% 
  group_by(name, month) %>% 
  summarise(across(prcp:tmin, ~ mean(.x, na.rm = TRUE)))
```

This is a dataframe!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_path()

weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 2) # get a knitter table in html
```


## Grouped mutates

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
    ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

lagged observations

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    yesterday_tmax = lag(tmax), # by default, lag by 1; the inverse function of `lag()` is `lead()`
    tmax_change = tmax - yesterday_tmax
  ) %>% 
  summarise(
    sd_tmax_change = sd(tmax_change, na.rm = TRUE)
  )
```


one other window function ...

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    tmax_rank = min_rank(tmax) # dealing with the ties
  )

weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    tmax_rank = min_rank(desc(tmax)) # the warmest day ranks as 1
  )

weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    tmax_rank = min_rank(tmax)
  ) %>% 
  filter(tmax_rank < 4) %>% 
  arrange(name, month, tmax_rank)
```

