---
title: "viz_part_01"
author: "Siyue Gao"
date: "2022-09-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data

```{r}
library(tidyverse)
library(ggridges)
```

This code chunk from the course website gets the data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Let's make a scatterplot: minimum temperature vs. maximum temperature.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Let's make the same scatterplot but different.

```{r}
weather_df %>%
  drop_na() %>% 
  filter(name == "CentralPark_NY") %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Let's keep making the same plot but different.

```{r}
weather_scatterplot =
  weather_df %>%
  drop_na() %>%
  ggplot(aes(x = tmin, y = tmax))

weather_scatterplot +
  geom_point()
```

## Let's fancy this up a bit!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name))

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + # define color only in the scatterplot
  geom_smooth() # only 1 curve

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + # applying color both to scatterplot and smooth lines
  geom_point() +
  geom_smooth()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + # define color only in the scatterplot
  geom_smooth(aes(color = name))

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + # applying color both to scatterplot and smooth lines
  geom_point(alpha = .3) + # making the points more or less transparent (0, 1)
  geom_smooth(se = FALSE) # getting rid of the error bars

# Not recommended - Always include the data as you can!
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + # applying color both to scatterplot and smooth lines
  geom_smooth(se = FALSE)
```

Mayber make separate panels.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) # row ~ column (. is just a placeholder)

# can tell central park has a much wider range of temperature

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) + 
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

`tmax` vs `tmin` is boring, let's spice it up some.

Maybe seasonal trends of maximum temperature?

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Some quick stuff

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

